Assessment and implementation of resource road simulation methods for projection of cumulative effects of disturbance on wildlife

### 2023-04-13 analysis/paper/paper.Rmd
reconciling.

## Aspatial performance
For each projection method variant the aspatial performance varied across the five road network metrics and also varied depending on whether we consider scores within cutblocks, outside cutblocks, or overall (Figure \@ref(fig:aspat-perf-fig)). Within all cutblocks, the Hardy method most accurately projected road density, and least accurately projected distance to road. Low density method variants most severely underprojected road density within cutblocks, and most accurately projected distance to road within cutblocks. These apparently contradictory results are explained by areas in our input data that have been misidentified as cutblocks with no associated observed roads (e.g. Figure \@ref(fig:proj-map-fig)b). Omitting cutblocks without observed roads from the comparison (Figure \@ref(fig:aspat-perf-fig) row 2) clarifies that within cutblocks the Hardy method most accurately projected road density and distance to road. However, note that our observed landscape is densely roaded in part because switchbacks are required in steep terrain (Figure \@ref(fig:proj-meth-fig)). The Hardy method projects more wiggles than the other methods, but none of the methods considered here can project switchbacks (Figure \@ref(fig:proj-meth-fig)).

@@

```{r get-data}
# load csv
meanTable <- read.csv(here(data_path_drvd, "mean_table.csv"))
meanTableRealCuts <- read.csv(here(data_path_drvd,"../TSA27_real_cuts",
                                   "mean_table.csv"))
matchData <- read.csv(here(data_path_drvd, "agree_table.csv"))

@@

- as expected, none of these algorithms are great at projecting exact road locations. we selected revelstoke because it is topographically complex, so a good challenge. Hardy result in eastern Canadian shield shows better overall accuracy projecting road density in less extreme terrain when the costs of water crossings are more carefully accounted for. Stream crossing behaviour in Revelstoke landscape could improved with more a detailed high resolution cost surface. But our results highlight that performance of these algorithms depends not only on accounting for cost but on complexity of the terrain. In particular, none of these simple algorithms will project switchbacks on steep slopes. A related fundamental limitation is that none of these algorithms can project tortuosity within raster cells, so in complex landscapes projections of road density will become less accurate as raster cell resolution increases. 
- Aspatial overall performance metrics can be difficult to interpret. To more deeply understand variation in projection accuracy we found it helpful to distinguish between performance within cutblocks and outside of them, and to examine spatial variation in discrepancies. Examination of spatial mismatch clarified that some discrepancies in our landscape are due to errors in our input data, rather than projection errors. Better harvest & existing road data would allow us to distinguish errors in projections from errors in the input data (cite lit in coops paper)

>>>>

## Aspatial performance
For each projection method variant the aspatial performance varied across the five road network metrics and also varied depending on whether we consider scores within cutblocks, outside cutblocks, or overall (Figure \@ref(fig:aspat-perf-fig)). Within all mapped cutblocks, the Hardy method most accurately projected road density, and least accurately projected distance to road. Low density method variants most severely underprojected road density within cutblocks, and most accurately projected distance to road within cutblocks. These apparently contradictory results are explained by areas in our input data that have been misidentified as cutblocks with no associated observed roads (e.g. Figure \@ref(fig:proj-map-fig)b). Omitting cutblocks without observed roads from the analysis (Figure \@ref(fig:aspat-perf-fig)) clarifies that within real cutblocks the Hardy method most accurately projected road density and distance to road. However, note that our observed landscape is densely roaded in part because switchbacks are required in steep terrain (Figure \@ref(fig:proj-meth-fig)). The Hardy method projects more tortuous roads within cutblocks, but none of the methods considered here can project switchbacks (Figure \@ref(fig:proj-meth-fig)).

@@

```{r get-data}
# load csv
meanTable <- read.csv(here(data_path_drvd, "mean_table.csv"))
#meanTableRealCuts <- read.csv(here(data_path_drvd,"../TSA27_real_cuts",
#                                   "mean_table.csv"))
matchData <- read.csv(here(data_path_drvd, "agree_table.csv"))

@@

- we selected revelstoke because it is topographically complex, so a good challenge. Hardy result in eastern Canadian shield showed better overall accuracy projecting road density in less extreme terrain when the costs of water crossings are more carefully accounted for. Stream crossing behaviour in Revelstoke landscape could be improved with more a detailed high resolution cost surface. But our results highlight that performance of these algorithms depends not only on accounting for cost but on complexity of the terrain. In particular, none of these simple algorithms will project switchbacks on steep slopes. A related fundamental limitation is that none of these algorithms can project tortuosity within raster cells, so in complex landscapes projections of road density will become less accurate as raster cell resolution increases. 
- Aspatial overall performance metrics can be difficult to interpret. To more deeply understand variation in projection accuracy we found it helpful to distinguish between performance within cutblocks and outside of them, and to examine spatial variation in discrepancies. Examination of spatial mismatch clarified that some discrepancies in our landscape are due to errors in our input data, rather than projection errors, and removing mapped cutblocks without observed access roads from our analysis helped resolve the apparently contradictory result finding that me Better harvest & existing road data would allow us to more distinguish errors in projections from errors in the input data (cite lit in coops paper)
